---
title: "Non-immune celltype annotation on GBM/HGG Xenium data"
author: "heinin"
date: "2025-10-11"
output:
  workflowr::wflow_html:
    code_folding: hide
  html_notebook: default
editor_options:
  chunk_output_type: console
---

## Import packages

```{r, warning=F, message=F}

suppressPackageStartupMessages({
  library(workflowr)
  library(arrow)
  library(Seurat)
  library(SeuratObject)
  library(SeuratDisk)
  library(tidyverse)
  library(tibble)
  library(ggplot2)
  library(ggpubr)
  library(ggrepel)
  library(googlesheets4)
  library(workflowr)})

```

## Environment variables and helper functions

```{r, message = F, warning = F}

setwd("/home/hnatri/13384_GBMHGG_SPP1_Xenium/")
set.seed(9999)
options(scipen = 99999)
options(ggrepel.max.overlaps = Inf)

source("/home/hnatri/13384_GBMHGG_SPP1_Xenium/code/colors.R")
source("/home/hnatri/13384_GBMHGG_SPP1_Xenium/code/plot_functions.R")

```

## Feature annotations

```{r, message = F, warning = F, fig.width = 4, fig.height = 7}

gs4_deauth()
all_sheets  <- gs4_get("https://docs.google.com/spreadsheets/d/1V1YxYPMI-J1ilZ3ooTDM323s5VMIf7E9f5XsuIOaGAg/edit?usp=sharing")
sheet_names(all_sheets)
multissue_cancer <- read_sheet(all_sheets, sheet = "Multi-tissue and cancer basepanel")
custom_panel <- read_sheet(all_sheets, sheet = "Custom probes")

immune_annot <- read_sheet(all_sheets, sheet = "Immune annotations")

```

## Annotation groups

```{r, message = F, warning = F, fig.width = 4, fig.height = 7}

# Markers
main_plot_features <- c("PTPRC", "CD3D", "CD3E", "CD4", "CD8A", "JAK1", "JAK2", "STAT4", "STAT3", "TIGIT", "GZMB", "SELL", "CD19", "CD68", "CD44", "MARCO", "APOE", "C1QB", "C1QBP", "CD163", "SPP1", "VCAN", "MUC5AC", "NOTCH3", "RGS5", "MS4A1", "PGA5", "PLVAP", "CCL21", "LYVE", "SPARCL1", "FN1", "DCN", "LUM", "COL1A2", "PDPN", "ACTA2", "LMNA", "EGR3", "TP53", "JUN", "FOS", "BRAF", "KIT", "SOX9", "RNF43", "EPCAM", "PECAM", "CEACAM", "TGFB1", "JUNB", "TCF7", "GZMA", "GZMK", "CD47", "CD74", "LYZ", "TREM2", "CD247", "COL5A1", "SPARC", "CD36", "CTLA4", "RGS1", "CCL2", "SELENOP", "LYVE1", "CTSD", "MMP9", "CDK1", "FLT1", "RAMP2", "CCL14", "MCAM", "NRP1", "CEACAM6", "IL32", "TGFBI", "VEGFA", "LAMP1", "IGFBP7", "POSTN", "C1S", "C1R", "CSF1", "NOTCH1", "FOXP3", "IL2RB", "TUBA1B", "S100A9", "C3", "CXCR2", "SDC1", "CD38", "IGHG1", "PRDX4", "ICAM1", "GATA2", "CXCL3")

neuro <- c("OLIG1", "OLIG2", "AQP4", "APOLD1", "CCL4", "RELN", "PLP1", "PDGFRA")
mycafs <- c("TGFB1", "ACTA2", "TAGLN") #"HOPX", "MYL9"
icafs <- c("IL6", "CXCL12", "CXCL14") #"CXCL13", 
abcafs <- c("CD74", "HLA-DPB1", "HLA-DPB2", "HLA-DQA1", "HLA-DRA", "HLA-DRB1")
ficafs <- c("SFRP2", "SFRP4", "COMP", "CXCL14", "PLA2G2A", "ASPN")

main_plot_features <- c(main_plot_features, mycafs, icafs, abcafs, ficafs, neuro)
main_plot_features <- unique(main_plot_features)

# Smaller set for feature plots
plot_features <- c("PTPRC", "CD3D", "CD3E", "CD4", "CD8A", "JAK1", "STAT3", "TIGIT", "GZMB", "SELL", "CD19", "CD68", "CD44", "MARCO", "APOE", "C1QB", "C1QBP", "CD163", "SPP1", "LYVE", "SPARCL1", "FN1", "DCN", "LUM", "COL1A2", "PDPN", "EPCAM", "PECAM", "CEACAM", "TGFB1", "JUNB", "TCF7", "GZMA", "GZMK", "CD47", "CD74", "LYZ", "TREM2", "CD247", "COL5A1", "SPARC", "CD36", "CTLA4", "RGS1", "CCL2", "SELENOP", "LYVE1", "FOXP3", "IL2RB", "TUBA1B", "S100A9", "SDC1")

fibroblast_markers <- multissue_cancer[grep("Fibroblast", multissue_cancer$Annotation),]$Gene # 65

immune_markers <- multissue_cancer[grep("Immune", multissue_cancer$Annotation),]$Gene # 6

macrophage_markers <- multissue_cancer[grep("Macrophage", multissue_cancer$Annotation),]$Gene # 70
granulocyte_markers <- multissue_cancer[grep("Granulocytes", multissue_cancer$Annotation),]$Gene # 25
dendritic_markers <- c(multissue_cancer[grep("dendritic", multissue_cancer$Annotation),]$Gene,
                       multissue_cancer[grep("Dendritic", multissue_cancer$Annotation),]$Gene) #22

tcell_markers <- multissue_cancer[grep("T cells", multissue_cancer$Annotation),]$Gene # 62
bcell_markers <- multissue_cancer[grep("B cells", multissue_cancer$Annotation),]$Gene # 45

proliferation_markers <- multissue_cancer[grep("Proliferation", multissue_cancer$Annotation),]$Gene # 7
endothelial_markers <- multissue_cancer[grep("Endothelial", multissue_cancer$Annotation),]$Gene # 82

```

## Import data

```{r, message = F, warning = F, fig.width = 4, fig.height = 4}

# Copied to isilon /tgen_labs/banovich/PIPAC/Seurat
seurat_data <- readRDS("/tgen_labs/banovich/BCTCSF/13384_GBMHGG_Xenium/Seurat/nonimmune_clustered_NN30_PC20_Seurat.rds")

seurat_data <- NormalizeData(seurat_data)
seurat_data <- ScaleData(seurat_data)

# saveRDS(seurat_data, "/tgen_labs/banovich/BCTCSF/13384_GBMHGG_Xenium/Seurat/cell_clustered_NN30_PC50_Seurat.rds")

DimPlot(seurat_data,
        group.by = "leiden_1.5",
        cols = nonimmune_cluster_col,
        reduction = "umap",
        raster = T,
        label = T) +
  coord_fixed(ratio = 1) +
  theme_classic() +
  NoLegend()

```

```{r, message = F, warning = F, fig.width = 4, fig.height = 7}

DimPlot(seurat_data,
        reduction = "sp",
        group.by = "leiden_1.5",
        cols = nonimmune_cluster_col,
        raster = T,
        label = F) +
    coord_fixed(ratio = 1) +
    theme_minimal() +
    NoLegend() +
    ggtitle("TMA1_0069754")

```

### Feature expression

```{r, message = F, warning = F, fig.width = 10, fig.height = 20}

DotPlot(seurat_data,
        group.by = "leiden_1.5",
        features = main_plot_features,
        cols = c("gray89", "tomato3"))

```

```{r, message = F, warning = F, fig.width = 12, fig.height = 14}

FeaturePlot(seurat_data,
            slot = "data",
            features = plot_features,
            order = T,
            ncol = 5,
            reduction = "umap",
            raster = T,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

```


```{r, message = F, warning = F, fig.width = 12, fig.height = 30}

FeaturePlot(seurat_data,
            slot = "data",
            #split.by = "Sample",
            features = plot_features,
            order = T,
            reduction = "sp",
            raster = T,
            ncol = 4,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

```

### DenoIST

```{r, message = F, warning = F, fig.width = 10, fig.height = 4, eval = F}

DefaultAssay(seurat_data) <- "denoist_RNA"
DotPlot(seurat_data,
        group.by = "leiden_1.5",
        features = plot_features,
        cols = c("azure", "tomato3")) +
  RotatedAxis()

```

### Top cluster markers

```{r, message = F, warning = F, fig.width = 5, fig.height = 4}

Idents(seurat_data) <- seurat_data$leiden_1.5
cluster_markers <- FindAllMarkers(seurat_data,
                                  return.thresh = 0.01,
                                  logfc.threshold = 0.5,
                                  min.pct = 0.20,
                                  verbose = T)

table(cluster_markers$cluster)

output_cluster_markers <- cluster_markers %>%
  arrange(dplyr::desc(avg_log2FC)) %>%
  group_by(cluster) %>%
  dplyr::slice(1:30)

write.table(output_cluster_markers, "/home/hnatri/13384_GBMHGG_SPP1_Xenium/nonimmune_top_markers.tsv",
            quote = F, row.names = F, sep = "\t")

#hist(cluster_markers$avg_log2FC, main = "", xlab = "avg_log2FC", breaks = 100)
#hist(cluster_markers$p_val, main = "", xlab = "p_val", breaks = 100)
#hist(cluster_markers$p_val_adj, main = "", xlab = "p_val_adj", breaks = 100)

top_cluster_markers <- cluster_markers %>%
  arrange(dplyr::desc(avg_log2FC)) %>%
  group_by(cluster) %>%
  dplyr::slice(1:5)

```


```{r, message = F, warning = F, fig.width = 8, fig.height = 18, eval = T}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = unique(top_cluster_markers$gene),
                       group_var = "leiden_1.5",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

Canonical markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 14, eval = T}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = unique(main_plot_features),
                       group_var = "leiden_1.5",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

## Features by group

### Fibroblast

```{r, message = F, warning = F, fig.width = 8, fig.height = 12}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = fibroblast_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### Immune markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 4}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = immune_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### Macrophage markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 12}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = macrophage_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### Granulocyte markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 8}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = granulocyte_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### Dendritic markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 7}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = dendritic_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### T cell markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 12}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = tcell_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### B cell markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 14}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = bcell_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### Proliferation cell markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 4}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = proliferation_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### Endothelial cell markers

```{r, message = F, warning = F, fig.width = 8, fig.height = 14}

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = endothelial_markers,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

### Custom features

```{r, message = F, warning = F, fig.width = 8, fig.height = 26, eval = F}

all_genes <- c(custom_panel$Gene) #multissue_cancer$Gene
all_genes <- gsub("CD11b", "ITGAM", all_genes)
all_genes <- gsub("CD11c", "ITGAX", all_genes)
all_genes <- gsub("CD31", "PECAM1", all_genes)
all_genes <- gsub("CD45", "PTPRC", all_genes)
all_genes <- gsub("HLA-DR", "HLA-DRA", all_genes)
all_genes <- gsub("HLA.DPB1", "HLA-DPB1", all_genes)
all_genes <- gsub("IFNGR", "IFNGR1", all_genes)
all_genes <- gsub("TGFβ", "TGFB1", all_genes)
all_genes <- gsub("TGFβR", "TGFB1R", all_genes)
all_genes <- gsub("TGFB1R", "TGFBR1", all_genes)
all_genes <- gsub("IL-1R", "IL1R2", all_genes)

all_genes <- sort(all_genes)

create_dotplot_heatmap(seurat_object = seurat_data,
                       plot_features = all_genes,
                       group_var = "leiden_1.0",
                       group_colors = nonimmune_cluster_col,
                       column_title = "",
                       row_km = 5,
                       col_km = 5,
                       row.order = NULL,
                       col.order = NULL)

```

## Adding preliminary annotations

```{r, message = F, warning = F, fig.width = 8, fig.height = 26, eval = F}

```

## Reclustering ambiguous clusters

```{r, message = F, warning = F, fig.width = 8, fig.height = 26, eval = F}

```


